
\chapter{Optimizations for Scaling-up Materialized Knowledge Graph Construction Techniques}
\label{chapter:construction}


In this Chapter we introduce two approaches to optimize the construction of materialized knowledge graphs. Diverse approaches have been proposed to define the process of integrating heterogeneous datasets into knowledge graphs \citep{chebotko2009semantics,calvanese2017ontop,chaves2019what,priyatna2014formalisation}. Mapping languages (e.g., R2RML \citep{R2RML} and RML \citep{dimou2014rml}) and engines (e.g., RMLMapper\footnote{\url{https://github.com/RMLio/rmlmapper-java}} and RocketRML \citep{csimcsek2019rocketrml}) represent valuable contributions for performing this transformation process. Albeit highly used, existing approaches lack efficient data management techniques demanded to create knowledge graphs from large and heterogeneous datasets with duplicates. In the same manner as the approaches proposed in Chapter \ref{chapter:virtual}, the two contributions described exploit the information from mapping rules and ideas presented in Chapter \ref{chapter:mappig-translation} to scale-up the construction of KG.

Section \ref{chap7_rdfizer} presents an efficient set of data structures and their corresponding physical operator to scaling up the construction of materialized KGs where the mappings are following the RML specification. We empirically demonstrate the efficiency of the proposal and compare the obtained results against state of the art RML engines such as RocketRML y RMLMapper. Section \ref{chap7_funmap} describes FunMap, a set of optimization techniques for efficiently pre-processing functional mappings (mappings such as RML+FnO, that include transformation functions) and generation of function-free RML mappings that can be used by any RML parser. The two contributions presented in this Chapter are joint collaborations with the Scientific Data Management Group from TIB research center.



\input{7_kgconstruction/rdfizer}

\input{7_kgconstruction/funmap}

